---
title: "Bar_Chart"
output: pdf_document
date: "2025-12-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library and data
```{r}

getwd() ## get the working directory

## Load libraries
suppressMessages(library(tidyverse))
library(lattice)
library(palmerpenguins)

## Load data
school_front <- read.csv("Job Applicants DATA 12-29.csv") ## recruitment info.
View(school_front)

```

## Summary
```{r}

school_front <- as_tibble(school_front) ## convert to structured data frame

job_title <- school_front$Job.Title

summarize(school_front) ## the summary

```

## The first two arguments to ggplot() are data and mapping, and the first two arguments to aes() are x and y.

```{r}

## Job Titles Distribution
bar_plot <- ggplot(school_front) +
  geom_bar(mapping = aes(x = job_title))

school_front %>% 
  count(job_title)

## Certification Distribution
ggplot(school_front) +
  geom_bar(mapping = aes(x = Cert.Area.List))

school_front %>% 
  count(Cert.Area.List)

## Civil Service Test Distribution
ggplot(school_front) +
  geom_bar(mapping = aes(x = Civil.Service.Test))

school_front %>% 
  count(Civil.Service.Test) ## There are NA; more people said 'no' to civil service test (inference is that less people are taking civil service test)

## School District Distribution
ggplot(school_front) +
  geom_bar(mapping = aes(x = Currently.Employed.District))

school_front %>% 
  count(Currently.Employed.District)

## Law Conviction Distribution
ggplot(school_front) +
  geom_bar(mapping = aes(x = Law.Conviction, na.rm = TRUE))

school_front %>% 
  count(Law.Conviction) ## NA category is there; n = 2508 said 'no'; n = 92 said 'yes'

## Job Type Distribution
ggplot(school_front) +
  geom_bar(mapping = aes(x = Teacher...Admin.Positions))

school_front %>% 
  count(Teacher...Admin.Positions) ## Top positions; Administrator = 611; Teaching Assistant = 481

## Full-time or Part-time
ggplot(school_front) +
  geom_bar(mapping = aes(x = Employment.Status.Interests))

school_front %>% 
  count(Employment.Status.Interests) ## Re-factor these so it's just full-time, part-time, and other...

school_front$Areas.of.Proficiency ## Needs to be re-factored

school_front$Types.of.Positions ## Several categories could be combined

school_front$Application.Type.Statuses ## IDK yet

## Demographics

## Applicant City Distribution 
ggplot(school_front) +
  geom_bar(mapping = aes(x = Permanent.City))

school_front %>% 
  count(Permanent.City) ## This data has NA (" "); the data is not considering as NA though, empty string isn't the same

school_front$Permanent.State

school_front$How.did.you.hear.about.our.district...job. ## Determine which platforms are most effective

```

## Distribution Check (Continuous = Histogram)
```{r}

dates <- structured_data$Job.Deadline

years_experience <- school_front_data$Private.Years

ggplot(structured_data) +
  geom_histogram(mapping = aes(x = years_experience))


## ggplot(structured_data) +
  ## geom_histogram(mapping = aes(x = dates, binwidth = 0.5))

## Numerics
GPA <- na.omit(school_front$Under.Graduate.All.Courses.GPA)

```

## Using binwidth (jobs that have few vs. many candidates)
```{r}

small_subgroups <- structured_data %>% 
  filter(Candidates.Total.Count < 5)

var(small_subgroups) ## Least popular jobs

medium_subgroups <- structured_data %>% 
  filter(Candidates.Total.Count < 10)

medium_subgroups


large_subgroups <- structured_data %>% 
  filter(Candidates.Total.Count > 15)

large_subgroups ## Most popular jobs

```

## Clustering Effects (reveal sub-groups)
```{r}

## ggplot(data = small_subgroups, mapping = aes(x = )) +
  ## geom_histogram(binwidth = 0.01)

```

## Replacing the unusual values with missing values
```{r}

## You can use the ifelse() function to replace unusual values with NA.

diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y)) ## use mutate() to replace the variable with a modified copy

```

## Boxplots (identify outliers) ## 130 - 150k has the most outliers...
```{r}

box_plots <- ggplot(data = structured_data, mapping = aes(x = Candidates.Total.Count, y = Pay.Scale)) +
  geom_boxplot()

```

## Variation among classes
```{r}

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()

```

## Scatterplot Distribution
```{r}

scatter_plot <- ggplot(data = structured_data) + 
  geom_point(mapping = aes(x = Candidates.Total.Count, y = Pay.Scale))

scatter_plot

```

## Parsing
```{r}

pay.scale <- parse_number(structured_data$Pay.Scale)

## typeof(pay.scale)
## sd(pay.scale)
## var(pay.scale)

```

## Relational Data
```{r}

## nycflights13 contains four tibbles that are related to the flights table

## top_dest <- flights %>%
  ## count(dest, sort = TRUE) %>%
  ## head(10)
## top_dest

## flights %>% 
  ## filter(dest %in% top_dest$dest)

```

## Replacement
```{r}

## With str_replace_all() you can perform multiple replacements by supplying a named vector

x <- c("Per contract", 
       "Based on experience and demonstrated abilities",
       "Negotiable")

str_replace_all(x, c("Per contract" = "Not a number", "Based on experience and demonstrated abilities" = "NA", "Negotiable" = "This still isn't a number"))

```

## Replacement (test)
```{r}

x <- structured_data$Pay.Scale

str_replace_all(x, c("Per contract" = "Not a number",
                     "Based on experience and demonstrated abilities" = "NA",
                     "Negotiable" = "This still isn't a number"))

```

## Arrange the data
```{r}

date <- structured_data$Tracker.Date

title <- structured_data$Job.Title

arrange(structured_data, date)

## Numeric Variables
years_experience <- wincap_data$Private.Years
GPA <- wincap_data$Under.Graduate.All.Courses.GPA
service_years <- school_front$Public.Years

cor.test(service_years, years_experience) ##




```

```{r}

sample_data <- penguins

View(sample_data)

cor.test(penguins$bill_length_mm, penguins$flipper_length_mm, method = "pearson") ## Moderate correlation

```

```{r}

ggplot(na.omit(penguins), 
       # aesthetic mappings
       aes(
         # flipper length (y) vs body mass (x)
         x = body_mass_g, y = flipper_length_mm, 
         # color AND shape of points vary by island
         color = island, shape = island)) +
  # geom_point to add points layer
  geom_point() + 
  # specify labels for ALL mappings (both Island labels required)
  labs(x = "Body size (g)", 
       y = "Flipper length (mm)",
       colour = "Island", shape = "Island") +
  # facet on sex by species grid
  facet_grid(rows = vars(sex), cols = vars(species))

```

## Colors
```{r}

 set.seed(19)
 x <- rnorm(30)
 y <- rnorm(30)
 plot(x, y, col = rep(1:3, each = 10), pch = 19)
 legend("bottomright", legend = paste("Group", 1:3), col = 1:3, pch = 19, bty = "n")

```

```{r}

penguins %>% 
  group_by(species) %>% 
  summarise(mean_bill_length = mean(bill_length_mm, na.rm = TRUE))

```

```{r}

sorting <- select(structured_data, starts_with("Newark"))
sorting

```








