---
title: "Stat Models"
output: pdf_document
date: "2025-12-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

suppressMessages(library(tidyverse))
library(modelr)
library(gapminder)

```

## A simple model (linear)
```{r}

ggplot(sim1, aes(x, y)) + 
  geom_point()

## The Model
models <- tibble(
  a1 = runif(250, -20, 40), ## Run 250 models
  a2 = runif(250, -5, 5)
)

ggplot(sim1, aes(x, y)) + 
  geom_abline(aes(intercept = a1, slope = a2), data = models, alpha = 1/4) +
  geom_point() 

```

## Finding Residuals (This distance is the prediction and the actual y value in the data)
```{r}

model1 <- function(a, data) {
  a[1] + data$x * a[2]
}
model1(c(7, 1.5), sim1)

```

## Root square mean for ANOVA
```{r}

measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff ^ 2))
}
measure_distance(c(7, 1.5), sim1)


```
## Use purrr to compute the distance for all the models defined above.
```{r}

sim1_dist <- function(a1, a2) {
  measure_distance(c(a1, a2), sim1)
}

models <- models %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))
models

```

## Best Model means smallest distance between predicted and observed
```{r}

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(models, rank(dist) <= 10)
  )

```

## We could represent as a scatterplot (with loss of 0 being the best)
```{r}

ggplot(models, aes(a1, a2)) +
  geom_point(data = filter(models, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist))

```
## Conducting a grid search
```{r}

grid <- expand.grid(
  a1 = seq(-5, 20, length = 25),
  a2 = seq(1, 3, length = 25)
  ) %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))

grid %>% 
  ggplot(aes(a1, a2)) +
  geom_point(data = filter(grid, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist)) 

```

## How to choose which model
```{r}

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(grid, rank(dist) <= 10)
  )

```

## Iteratively making the grid finer and finer until you narrowed in on the best model. 
```{r}

best <- optim(c(0, 0), measure_distance, data = sim1)
best$par ## you pick a starting point and look around for the steepest slope.

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])

```

## Fitting a linear model
```{r}

sim1_mod <- lm(y ~ x, data = sim1)
coef(sim1_mod)

```
## First argument is a data frame, and for each subsequent argument it finds the unique variables and then generates all combinations
```{r}

grid <- sim1 %>% 
  data_grid(x) 
grid

grid <- grid %>% 
  add_predictions(sim1_mod) ## add the predictions
grid

```

## Add the plot line (predictions)
```{r}

ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred), data = grid, colour = "red", size = 1)

```

```{r}

sim1 <- sim1 %>% 
  add_residuals(sim1_mod)
sim1

ggplot(sim1, aes(resid)) + 
  geom_freqpoly(binwidth = 0.5)

```

## One-hot encoding (converting categorical to continuous using binary)
```{r}

df <- tribble(
  ~ sex, ~ response,
  "male", 1,
  "female", 2,
  "male", 1
)
model_matrix(df, response ~ sex)

```

## Categorical predictor and continuous predictor
```{r}

ggplot(sim3, aes(x1, y)) + 
  geom_point(aes(colour = x2))

mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3) ## interaction

```

## Interactions (two continuous)
```{r}

mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)

grid <- sim4 %>% 
  data_grid(
    x1 = seq_range(x1, 5), 
    x2 = seq_range(x2, 5) 
  ) %>% 
  gather_predictions(mod1, mod2)
grid

```

## Graphing
```{r}

gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)
    
```

## Fit the model
```{r}

country_model <- function(df) {
  lm(lifeExp ~ year, data = df)
}

```

## Model Quality
```{r}

nz <- filter(gapminder, country == "New Zealand")
nz %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ")

nz_mod <- lm(lifeExp ~ year, data = nz)
nz %>% 
  add_predictions(nz_mod) %>%
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear trend + ")

nz %>% 
  add_residuals(nz_mod) %>% 
  ggplot(aes(year, resid)) + 
  geom_hline(yintercept = 0, colour = "white", size = 3) + 
  geom_line() + 
  ggtitle("Remaining pattern")

```

## Nested data
```{r}

by_country <- gapminder %>% 
  group_by(country, continent) %>% 
  nest()

by_country

## Load the data
gapminder

broom::glance(nz_mod)

models <- map(by_country$data, country_model)

by_country <- by_country %>% 
  mutate(model = map(data, country_model))
by_country


```

```{r}

by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)

```

## Evaluating Model Fit
```{r}

glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance, .drop = TRUE)

glance %>% 
  arrange(r.squared)

```

## Keeps columns together
```{r}

gapminder %>% 
  group_by(country, continent) %>% 
  nest()

```

## List-column - From vectorized functions
```{r}

df <- tribble(
  ~x1,
  "a,b,c", 
  "d,e,f,g"
) 

df %>% 
  mutate(x2 = stringr::str_split(x1, ",")) ## takes a character vector and returns a list of character vectors

```

```{r}

df %>% 
  mutate(x2 = stringr::str_split(x1, ",")) %>% 
  unnest(x2)

```







